The extended light profiles of the Sculptor and Ursa Minor dwarf galaxies: Nature or tidal nurture?

### 2025-09-29 thesis/introduction.tex
thesis edits

Dwarf galaxies have long raised conundrums for theories of galaxy
formation. The discovery of Fornax and Sculptor in 1938
\citep{shapley1938}\footnote{Technically, the Large and Small Magellanic
  Clouds (LMC, SMC) are also classified as dwarf galaxies, but these
  were likely always known to humans at southern latitudes.}, with no
known analogues, already presented such an enigma. H. Shapley presented
these dwarfs as a new type of \emph{stellar system} resembling the
Magellanic Clouds and globular clusters but did not attempt to speculate
on their exact nature. While dwarf galaxies were soon understood to be
galaxies based on the inferred luminosities and sizes, their nature
remained unclear for decades
\citep[e.g.,][]{hodge1971, gallagher+wyse1994}.

@@

Today, a common definition for a dwarf galaxy is a gravitationally bound
stellar system with dark matter.\footnote{Or, more generally, systems
  inconsistent with Newtonian dynamics of visible matter alone,
  \citet{willman+strader2012}} In contrast, star clusters (like GCs)
have no clear evidence for dark matter. The boundary between these two
classes blurs for faint, compact stellar associations, known as
``ambiguous'' systems.

Dwarf galaxies span a large range of sizes, luminosities, and
morphologies. Broadly, there are three classes of dwarf galaxies based
on luminosity. Local \textbf{bright dwarfs} with magnitudes
\(-14 \gtrsim M_V \gtrsim  -18\) or stellar masses
\(3\times10^7\,\Mo \lesssim M_\star \lesssim 10^9\,\Mo\) \footnote{assuming
  stellar mass-to-light ratio of 1, may be \(\sim 2\) for older
  populations}, often exhibit irregular morphologies and recent star
formation. Fig.~\ref{fig:galaxy_images} shows the Large Magellanic Cloud
(LMC) as an example of an irregular, bright dwarf galaxy, where most
stars are in a rotationally-supported thin disk (seen nearly face-on)
with a prominent bar. \textbf{Classical dwarf spheroidals}\footnote{While
  formally the dwarf galaxy names we discuss contain ``dwarf
  spheroidal'' (dSph), e.g.~Sculptor dSph, we omit this suffix for
  brevity. Additionally, the 12 classical dwarf satellites of our Galaxy
  are (in order of decreasing luminosity) Sagittarius, Fornax, Leo I,
  Sculptor, Antlia II, Leo II, Carina, Draco, Ursa Minor, Canes Venatici
  I, Sextans I, and Crater II. Antlia II, Crater II, and Canes Venatici
  I are the only post digital sky surveys additions.} occupy
intermediate luminosities ( \(-7.7 \gtrsim M_V  \gtrsim -14\) or
\(10^5\,\Mo \lesssim M_\star \lesssim 3\times10^7\,\Mo\)). Typically,
these systems are old, non-star forming, gas-poor, and spheroidal. All
Milky Way satellites discovered before digital sky surveys are
classicals, and these systems remain among the best studied. The
\textbf{ultra-faint}s dwarfs (UFDs) occupy the very faintest magnitude
range (\(M_V \gtrsim -7.7\) or \(M_\star \lesssim 10^5\,\Mo\)). These
galaxies have minuscule stellar masses, typically compact sizes, and
very metal-poor stellar populations \citep[see review][]{simon2019}.
Altogether, known dwarf galaxies span more than 15 orders in magnitude,
or over 7 decades in stellar mass.

Most well-studied dwarf galaxies lie near the Milky Way or in its
vicinity, the Local Group (LG) of galaxies. The LG is defined as the
group consisting of galaxies within \(\sim 1\) Mpc from the Milky
Way-Andromeda centre \citep[e.g.,][ and references
therein]{mcconnachie2012}. Today, we know that the Milky Way system is
teeming with dwarfs, many of which are satellites of either the Milky
Way or(MW) or Andromeda (M31). Fig.~\ref{fig:mw_satellite_system} shows
the MW satellite system, including dwarf galaxies, globular clusters,
and ambiguous systems. Such a nearby population of dwarf galaxies are
useful for resolved, detailed studies aiming to understand the nature of
these systems.

@@

The very early universe was almost featureless. Our earliest
observations of the universe stem from the cosmic microwave background
(CMB)---revealing a uniform, isotropic, near-perfect blackbody emission.
But tiny perturbations in the CMB, temperature fluctuations of 1 part in
10,000, reveal the underlying seeds of large-scale cosmological
structure. In an expanding universe, gravitational instability makes CDM
overdensities grow and collapse hierarchically onto larger structures.
Initially, baryonic matter was coupled to radiation and resisted
collapse. Dark matter, only influenced by gravity instead, freely
collapsed into the first structures. Mass perturbations sufficiently
small and overdense become self-gravitating structures, known as
\emph{halos}. After recombination, where electrons combined with atomic
nuclei to form atoms, baryons decoupled from radiation and fell into the
dark matter halos. The densest pockets of baryons later formed the first
stars and galaxies.

Dark matter halos, and their associated galaxies, rarely evolve in
isolation. Instead, \LCDM{} structure formation is \emph{hierarchical}.
Small dark matter halos collapse first and hierarchically merge into
progressively larger halos
\citep[e.g.,][]{blumenthal+1984, white+rees1978, white+frenk1991}.
Hierarchical assembly is evident through the large scale structure of
the universe, remnants of past mergers within the Milky Way, and tidal
disruption of dwarf galaxies and their streams around nearby galaxies.

Small-scale structure formation is sensitive to deviations from \LCDM{}
cosmology \citep[e.g.,][]{bechtol+2022}. One key prediction of \LCDM{}
is that mass perturbations are expected to exist on all scales, and are
largest on the smallest scales, so we would expect the formation of
halos on all scales. Many alternative models, such as warm dark matter,
may smooth out small-scale features and reduce the abundance of small
halos or change their structure \citep[e.g.,][]{lovell+2014}. Dwarf
galaxies, which occupy the smallest dark matter halos, are promising
windows into the behaviour of dark matter on small scales.

@@

In \LCDM{} cosmological simulations, dark matter halos are remarkably
self-similar. In \citet{NFW1996, NFW1997}, hereafter NFW, the authors
observe that the spherically-averaged density profiles \(\rho(r)\) are
universally well described by a two-parameter law,
\begin{equation}\protect\phantomsection\label{eq:nfw}{
\rho/\rho_s= \frac{1}{(r/r_s)(1+r/r_s)^2},
}\end{equation} where \(r_s\) is a scale radius and \(\rho_s\) a scale
density. This profile has shown remarkable success in describing \LCDM{}
halos across several orders of magnitude in mass. NFW profiles are
\emph{cuspy}, where the density rises like \(\rho \sim 1/r\) at small
radii \(r \ll r_s\). The steepness of the density profile increases
gradually with radius, and at large radii the density falls off like
\(\rho \sim 1/r^3\). The blue solid curve in Fig.~\ref{fig:nfw_density}
shows an example NFW halo.

The total mass of an NFW profile formally diverges, so halo masses are
conventionally defined using an overdensity criterion. The virial mass,
\(M_{200}\), is defined as the mass within a radius, \(r_{200}\),
containing a mean enclosed density 200 times\footnote{For the collapse
  of a uniform spherical density, the virialized overdensity would be
  \(\Delta = 18\pi^2\approx 178\) for a critical universe
  \(\Omega_m = 1\). This is commonly rounded to \(\Delta = 200\). While
  this parameter may be closer to \(\Delta \approx 100\) for our
  universe, \(\Delta\) also increases with redshift \citep[using eq. 6
  from][]{bryan+norman1998}.} the critical density of the universe:
\begin{equation}{
M_{200} =200\,\frac{4\pi}{3} \ r_{200}^3\ \rho_{\rm crit}, \qquad {\rm where} \quad \rho_{\rm crit}(z) = 3H(z)^2 / 8\pi G,
}\end{equation} and \(H(z)\) is the Hubble constant as a function of
redshift. Another way of characterizing NFW halos is through the halo
concentration, \(c=r_{200} / r_s\), which describes how the
characteristic radius scale of the halo compares to the virial radius.
Using this parameter, the scale density is a function of \(c\) alone,
\(\rho_s = (200/3)\,\rho_{\rm crit} c^3 / [\log(1+c) - c/(1+c)]\)
\citep{NFW1996}.

@@

The two parameters of an NFW profile are not independent. Lower-mass
dark matter halos often collapse earlier, when the universe was denser.
As a result, low mass subhalos tend to be more concentrated
\citep[e.g.,][]{NFW1997}. The relationship between \(M_{200}\) and c, or
the mass-concentration relation, describes the mean trend of
concentration with mass or, equivalently, the dependence of \(\vmax\) on
\(\rmax\) \citep[e.g.,][]{bullock+2001, ludlow+2016}. The left panel of
Fig.~\ref{fig:smhm} illustrates the present-day mass-concentration from
\citet{ludlow+2016}. While concentration tends to decrease with
increasing mass, the relation has substantial scatter. Other parameters
such as the halo spin or shape may affect the scatter of the
mass-concentration relation, but their effect is typically expected to
be small \citep{navarro+2010, dicintio+2013, dutton+maccio2014}. The
mass-concentration relation informs our assumed structure of Scl and UMi
later.

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{figures/example_density_profiles.png}
\caption[Example dark matter and stellar density profiles]{Density
profiles in log density versus log radius for stars and dark matter in a
Fornax-like galaxy. The dark matter is more extended and massive than
the star across the entire galaxy. The galaxy has a stellar mass
\(M_\star \approx 2.5\times10^7\,\Mo\) with half-light radius
\(0.65\,\kpc\) \citep{munoz+2018, woo+courteau+dekel2008}. The halo
represents the cosmological mean given the stellar mass, with
\(\vmax=40\,\kpc\) and \(\rmax=8\,\kpc\), or
\(M_{200} = 1\times10^{10}\,\Mo\) and
\(c=12.5\).}\label{fig:nfw_density}
\end{figure}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/cosmological_means.pdf}}
\caption[Cosmological halo mass and stellar mass
relations]{\textbf{Left} The NFW halo concentration \(c=r_{200} / r_s\)
as a function of virial mass \(M_{200}\). The solid line with
1-\(\sigma\) shaded region is the mass-concentration relation from
\citet{ludlow+2016} for \(z=0\). \textbf{Right} Stellar mass (top) as a
function of maximum circular velocity. The solid line with the
1-\(\sigma\) shaded region is the relation from \citet{fattahi+2018}
with scatter points simulated central galaxies from \apostle{} in
\citet{fattahi+2018}. The pink star illustrates the location of the
Fornax galaxy, whose density profiles are shown in
Fig.~\ref{fig:nfw_density}.}\label{fig:smhm}
\end{figure}

\subsection{\texorpdfstring{Galaxies formation in
\LCDM{}}{Galaxies formation in }}\label{galaxies-formation-in}

The observed abundance of galaxies may be compared with the abundance of
\LCDM{} halos to derive constraints regarding which galaxies inhabit
which halos. This technique, dubbed ``abundance matching,'' implies a
tight relation between the stellar mass of a galaxy and the mass of the
halo it inhabits \citep{li+white2009, moster+naab+white2013}.

>>>>

Dwarf galaxies have long raised conundrums for theories of galaxy
formation. The discovery of Fornax and Sculptor in 1938
\citep{shapley1938}\footnote{Technically, the Large and Small Magellanic
  Clouds (LMC, SMC) are also classified as dwarf galaxies, but these
  were likely always known to humans at southern latitudes.}, with no
known analogues at the time, already presented such an enigma. H.
Shapley presented these dwarfs as a new type of \emph{stellar system}
resembling the Magellanic Clouds and globular clusters but did not
attempt to speculate on their exact nature. While dwarf galaxies were
soon understood to be galaxies based on the inferred luminosities and
sizes, their nature remained unclear for decades
\citep[e.g.,][]{hodge1971, gallagher+wyse1994}.

@@

Today, a common definition for a dwarf galaxy is a gravitationally bound
stellar system with dark matter.\footnote{Or, more generally, systems
  inconsistent with Newtonian dynamics of visible matter alone,
  \citet{willman+strader2012}} In contrast, star clusters (like GCs)
have no clear evidence for dark matter. The boundary between these two
classes blurs for faint, compact stellar associations. Systems with
shared characteristics of globular clusters and dwarf galaxies, and
without evidence as to their nature, are known as ``ambiguous'' systems
(\textbf{ref}).

Dwarf galaxies span a large range of sizes, luminosities, and
morphologies. Broadly, there are three classes of dwarf galaxies based
on luminosity. Local \textbf{bright dwarfs} with magnitudes
\(-14 \gtrsim M_V \gtrsim  -18\) or stellar masses
\(3\times10^7\,\Mo \lesssim M_\star \lesssim 10^9\,\Mo\) \footnote{assuming
  stellar mass-to-light ratio of \(1 \Mo / L_{\odot}\), may be
  \(\sim 2 \Mo/L_\odot\) for older populations.}, often exhibit
irregular morphologies and recent star formation.
Fig.~\ref{fig:galaxy_images} shows the Large Magellanic Cloud (LMC) as
an example of an irregular, bright dwarf galaxy, where most stars are in
a rotationally-supported thin disk (seen nearly face-on) with a
prominent bar. \textbf{Classical dwarf spheroidals}\footnote{While
  formally the dwarf galaxy names we discuss contain ``dwarf
  spheroidal'' (dSph), e.g.~Sculptor dSph, we omit this suffix for
  brevity. Additionally, the 12 classical dwarf satellites of our Galaxy
  are (in order of decreasing luminosity) Sagittarius, Fornax, Leo I,
  Sculptor, Antlia II, Leo II, Carina, Draco, Ursa Minor, Canes Venatici
  I, Sextans I, and Crater II. Antlia II, Crater II, and Canes Venatici
  I are the only post digital sky surveys additions.} occupy
intermediate luminosities ( \(-7.7 \gtrsim M_V  \gtrsim -14\) or
\(10^5\,\Mo \lesssim M_\star \lesssim 3\times10^7\,\Mo\)). Typically,
these systems are old, non-star forming, gas-poor, and spheroidal. All
Milky Way satellites discovered before digital sky surveys are
classicals, and these systems remain among the best studied. The
\textbf{ultra-faint}s dwarfs (UFDs) occupy the very faintest magnitude
range (\(M_V \gtrsim -7.7\) or \(M_\star \lesssim 10^5\,\Mo\)). These
galaxies have minuscule stellar masses, typically compact sizes, and
very metal-poor stellar populations \citep[see the review
by][]{simon2019}. Altogether, known dwarf galaxies span more than 15
magnitudes, or over 7 decades in stellar mass.

Most well-studied dwarf galaxies lie near the Milky Way or in its
vicinity, the Local Group (LG) of galaxies. The LG is defined as the
group consisting of galaxies within \(\sim 1\) Mpc from the Milky
Way-Andromeda centre \citep[e.g.,][ and references
therein]{mcconnachie2012}. Today, we know that the Milky Way system is
teeming with dwarfs, many of which are satellites of either the Milky
Way (MW) or Andromeda (M31). Fig.~\ref{fig:mw_satellite_system} shows
the MW satellite system, including dwarf galaxies, globular clusters,
and ambiguous systems. This nearby population of dwarf galaxies are
opportune for resolved, detailed studies aimed to understand the nature
of these systems.

@@

The very early universe was almost featureless. Our earliest
observations of the universe stem from the cosmic microwave background
(CMB)---revealing a nearly uniform, isotropic blackbody emission. But
tiny perturbations in the CMB, temperature fluctuations of 1 part in
100,000, reveal the underlying seeds of large-scale cosmological
structure. In an expanding universe, gravitational instability makes CDM
overdensities grow and collapse hierarchically onto larger structures.
Initially, baryonic matter was coupled to radiation and resisted
collapse. Dark matter, only influenced by gravity instead, freely
collapsed into the first structures. Mass perturbations sufficiently
small and overdense become self-gravitating structures, known as
\emph{halos}. After recombination, where electrons combined with atomic
nuclei to form atoms, baryons decoupled from radiation and fell into
dark matter halos, where the condensed at the centre through radiative
energy losses. The densest pockets of baryons later formed the first
stars and galaxies.

Dark matter halos, and their associated galaxies, rarely evolve in
isolation. Instead, \LCDM{} structure formation is \emph{hierarchical}.
Small dark matter halos collapse first and hierarchically merge into
progressively larger halos
\citep[e.g.,][]{white+rees1978, blumenthal+1984, white+frenk1991}.
Hierarchical assembly is evident through the large scale structure of
the universe, remnants of past mergers within the Milky Way, and tidal
disruption of dwarf galaxies and their streams around nearby galaxies.

Small-scale structure formation is sensitive to deviations from \LCDM{}
cosmology \citep[e.g.,][]{bechtol+2022}. One key prediction of \LCDM{}
is that mass perturbations are expected to exist on all scales, and are
largest on the smallest scales, so we would expect the formation of
halos on all scales. Many alternative models, such as warm dark matter,
may smooth out small-scale features and reduce the abundance of small
halos or change their structure \citep[e.g.,][]{lovell+2014}. Dwarf
galaxies, which occupy the smallest dark matter halos, are promising
probes into the behaviour of dark matter on small scales.

@@

In \LCDM{} cosmological simulations, dark matter halos are remarkably
self-similar. In \citet{NFW1996, NFW1997}, hereafter NFW, the authors
observe that the spherically-averaged density profiles \(\rho(r)\) are
universally well described by a two-parameter law,
\begin{equation}\protect\phantomsection\label{eq:nfw}{
\rho/\rho_s= \frac{1}{(r/r_s)(1+r/r_s)^2},
}\end{equation} where \(r_s\) is a scale radius and \(\rho_s\) a scale
density. This profile has shown remarkable success at describing \LCDM{}
halos across several orders of magnitude in mass. NFW profiles are
\emph{cuspy}, where the density rises like \(\rho \sim 1/r\) at small
radii \(r \ll r_s\). The steepness of the density profile increases
gradually with radius, and at large radii the density falls off like
\(\rho \sim 1/r^3\). The blue solid curve in Fig.~\ref{fig:nfw_density}
shows an example NFW halo.

The total mass of an NFW profile formally diverges, so halo masses are
conventionally defined using an overdensity criterion. The virial mass,
\(M_{200}\), is defined as the mass within a radius, \(r_{200}\),
containing a mean enclosed density 200 times\footnote{For the collapse
  of a uniform spherical density, the virialized overdensity would be
  \(\Delta = 18\pi^2\approx 178\) for a critical universe
  \(\Omega_m = 1\). This is commonly rounded to \(\Delta = 200\). While
  this parameter may be closer to \(\Delta \approx 100\) for our
  universe, \(\Delta\) also increases with redshift \citep[using eq. 6
  from][]{bryan+norman1998}.} the critical density of the universe:
\begin{equation}{
M_{200} =200\,\frac{4\pi}{3} \ r_{200}^3\ \rho_{\rm crit}, \qquad {\rm where} \quad \rho_{\rm crit}(z) = 3H(z)^2 / 8\pi G,
}\end{equation} and \(H(z)\) is the Hubble constant, which depends on
redshift. Another way of characterizing NFW halos is through the
concentration parameter, \(c=r_{200} / r_s\), which describes how the
characteristic radial scale of the halo compares to the virial radius.
Using this parameter, the scale density is a function of \(c\) alone,
\(\rho_s = (200/3)\,\rho_{\rm crit} c^3 / [\log(1+c) - c/(1+c)]\)
\citep{NFW1996}.

@@

The two parameters of an NFW profile are not independent. Lower-mass
dark matter halos typically collapse earlier, when the universe was
denser. As a result, low mass subhalos tend to be more concentrated
\citep[e.g.,][]{NFW1997}. The relationship between \(M_{200}\) and c, or
the mass-concentration relation, describes the mean trend of
concentration with mass or, equivalently, the dependence of \(\vmax\) on
\(\rmax\) \citep[e.g.,][]{bullock+2001, ludlow+2016}. The left panel of
Fig.~\ref{fig:smhm} illustrates the present-day mass-concentration from
\citet{ludlow+2016}. While concentration tends to decrease with
increasing mass, the relation has substantial scatter. Other parameters
such as the halo spin or shape may affect the scatter of the
mass-concentration relation, but their effect is typically expected to
be small \citep{navarro+2010, dicintio+2013, dutton+maccio2014}.

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{figures/example_density_profiles.png}
\caption[Example dark matter and stellar density profiles]{Density
profiles in log 3D density versus log 3D radius for stars and dark
matter in a Fornax-like galaxy. The dark matter is more extended and
massive than the star across the entire galaxy. Fornax has a stellar
mass \(M_\star \approx 2.5\times10^7\,\Mo\) with half-light radius
\(0.65\,\kpc\) \citep{munoz+2018, woo+courteau+dekel2008}. The
corresponding cosmological-mean halo for Fornax has \(\vmax=40\,\kpc\)
and \(\rmax=8\,\kpc\), or \(M_{200} = 1\times10^{10}\,\Mo\) and
\(c=12.5\).}\label{fig:nfw_density}
\end{figure}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/cosmological_means.pdf}}
\caption[Cosmological halo mass and stellar mass
relations]{\textbf{Left} The NFW halo concentration \(c=r_{200} / r_s\)
as a function of virial mass \(M_{200}\). The solid line with
1-\(\sigma\) shaded region is the mass-concentration relation from
\citet{ludlow+2016} for \(z=0\). \textbf{Right} Stellar mass (top) as a
function of maximum circular velocity. The solid line with the
1-\(\sigma\) shaded region is the relation from \citet{fattahi+2018}
with scatter points simulated central galaxies from \apostle{} in
\citet{fattahi+2018}. The pink star illustrates the location of the
Fornax galaxy, whose density profiles are shown in
Fig.~\ref{fig:nfw_density}. \textbf{add rmax-vmax plot
too}}\label{fig:smhm}
\end{figure}

\subsection{\texorpdfstring{Galaxy formation in
\LCDM{}}{Galaxy formation in }}\label{galaxy-formation-in}

The observed abundance of galaxies may be compared with the abundance of
\LCDM{} halos to derive constraints regarding which galaxies inhabit
which halos. One simple technique, dubbed ``abundance matching,''
implies a tight relation between the stellar mass of a galaxy and the
mass of the halo it inhabits
\citep{li+white2009, moster+naab+white2013}.

### 2025-09-29 thesis/methods.tex
thesis edits

\begin{table*}[t]
\centering
\caption[Orbit initial conditions]{The orbital initial conditions for models presented in this work. The observables represent the medians from orbital integration used to derived the orbits. Instead, the initial position and velocity represents the initialization of the actual N-body model. The \smallperi{} represents instead the $3\sigma$ smallest pericentre, which we use to provide an upper limit on tidal effects. We describe the \texttt{LMC-flyby} orbit in Section \ref{sec:scl_lmc}. }
\label{tbl:orbit_ics}
\begin{tabular}{llll}
\toprule
Property & Scl: \smallperi{} & Scl: \verb|LMC-flyby| & Umi: \smallperi{}\\
\midrule
distance / kpc & 82.6 & 72.9 & 64.6\\
$\pmra / \masyr$ & 0.134 & 0.137 & -0.158\\
$\pmdec / \masyr$ & -0.198 & -0.157 & 0.050\\
LOS velocity / $\kms$ & 111.2 & 111.2 & -245.75\\
$t_i / \Gyr$ & -9.17 & -2.0 & -9.67\\
${x}_{i} / \kpc$ & -2.49 & 4.30 & 17.40\\
${y}_{i} / \kpc$ & -42.78 & 138.89 & 74.51\\
${z}_{i} / \kpc$ & 86.10 & 125.88 & 21.34\\
$\V_{x\,i} / \kms$ & -20.56 & 6.89 & 14.27\\
$\V_{y\,i} / \kms$ & -114.83 & -56.83 & 48.62,\\
$\V_{z\,i} / \kms$ & -57.29 & 52.09 & -114.08\\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Ursa Minor's Orbit}\label{ursa-minors-orbit}

>>>>

\begin{table*}[t]
\centering
\caption[Orbit initial conditions]{The orbital initial conditions for models presented in this work. The observables represent the medians from orbital integration used to derived the orbits. Instead, the initial position and velocity represents the initialization of the actual N-body model. The \smallperi{} represents instead the $3\sigma$ smallest pericentre, which we use to provide an upper limit on tidal effects. We describe the \texttt{LMC-flyby} orbit in Section \ref{sec:scl_lmc}. }
\label{tbl:orbit_ics}
\begin{tabular}{lllll}
\toprule
Property & Scl: \smallperi{} & Scl: \verb|LMC-flyby| & UMI: \smallperi{} & Umi: \smallperi{}.5\\
\midrule
distance / kpc & 82.6 & 72.9 & - & 64.6\\
$\pmra / \masyr$ & 0.134 & 0.137 & - & -0.158\\
$\pmdec / \masyr$ & -0.198 & -0.157 & - & 0.050\\
LOS velocity / $\kms$ & 111.2 & 111.2 & - & -245.75\\
$t_i / \Gyr$ & -9.17 & -2.0 & - & -9.67\\
${x}_{i} / \kpc$ & -2.49 & 4.30 &  & 17.40\\
${y}_{i} / \kpc$ & -42.78 & 138.89 &  & 74.51\\
${z}_{i} / \kpc$ & 86.10 & 125.88 &  & 21.34\\
$\V_{x\,i} / \kms$ & -20.56 & 6.89 &  & 14.27\\
$\V_{y\,i} / \kms$ & -114.83 & -56.83 &  & 48.62,\\
$\V_{z\,i} / \kms$ & -57.29 & 52.09 &  & -114.08\\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Ursa Minor's Orbit}\label{sec:orbit_corrections}
